### 2015-11-20 section/0_intro.tex
finish a pass of the whole paper

\item Root-cause taxonomy guided design. To provide the needed coverage, we follow
the root-cause taxonomy discussed above and design analysis routines for 
every root-cause sub-category. Given a candidate loop, we will 
apply a series of analysis to it to see if it matches any type of inefficiency.
%We will also put the analysis results from different analysis routines together
%to figure out the most likely root cause category for an inefficient loop.

>>>>

\item Taxonomy guided design. To provide good coverage, we follow
the root-cause taxonomy discussed above and design analysis routines for
every root-cause category. Given a candidate loop, \Tool
applies a series of analysis to it to see if it matches any type of inefficiency.
%We will also put the analysis results from different analysis routines together
%to figure out the most likely root cause category for an inefficient loop.

### 2015-11-20 section/1_under.tex
finish a pass of the whole paper

\subsubsection{Resultless loops}
A resultless loop spends a lot of time in computation without
producing any results useful after the loop.
It is further categorized based on which loop iterations are
(not) producing useful results.

@@

%\begin{figure}[h]
%\centering
%\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
%     morekeywords={+},keepspaces=true}
%  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/Mozilla347306.c}}
%\caption{A resultless 0*1? bug in Mozilla}
%\label{fig:Mozilla347306}
%\end{figure}

{\textbf{0*1?}}:
This type of loops only produce results in the last iteration, if any. 
They are often related to search: check a sequence of elements one
by one until the right one is found.
Clearly, whether this type of loops are efficient or not depends on the
workload. If they are indeed inefficient, they are often
fixed by data-structure changes.
As an example, Mozilla JavaScript engine used to use a linked list 
\texttt{script}
to store debugging information
for every bytecode instruction in a JavaScript file. 
Later on, people complained that Mozilla browsers were very slow in handling
large JavaScript files.
It turned out that those large
JavaScript files filled the linked list \texttt{script} with
tens of thousands of nodes. Repeated search through the
long \texttt{script} list then led to poor performance.
To fix this problem, 
developers simply replace linked list with hash table
(Mozilla Bug\#347306).
% This improve the performance by 52.60{\bf X}.

@@

\begin{figure}
\centering
\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/GCC46401.c}}
\caption{A resultless [0$|$1]* bug in GCC }
\label{fig:GCC46401}
\end{figure}

{\textbf{[0$|$1]*}}:
Each iteration in this type of loops may or may not produce results.
For some workload, the majority of iterations do not produce
results and lead to performance problems perceived by users.
These problems are often solved
by using a simple check to conditionally skip the expensive loop.
Figure~\ref{fig:GCC46401} shows such an example.
%how users perceive this problem
%explain how every iteration may or may not produce any results
%explain the fix
Users complained that compilation became extremely slow when they 
enabled the \texttt{-Wsequence-point} checking.
The slowness was caused
by the \texttt{for} loop shown in the figure. Since the algorithm
behind this loop is quadratic in the number of operands in an expression, 
programs with long expressions suffer severe slow-downs.
After further diagnosis, developers observed that this loop rarely
had any side-effects, as the \texttt{if} condition was rarely satisfied.
At the end, the patch greatly improved performance through a
simple checking that allows the loop to be
skipped for most expressions.

@@

{\textbf{[1]*}}:
Loops in this category always generate results in almost all iterations. 
They are inefficient because their results are useless due to
some high-level semantic reasons.
Understanding and fixing this type of inefficiency problems often require
deep understanding of the program and are difficult to automate.
For example, several Mozilla performance problems are caused by 
loops that contain intensive GUI operations whose graphical outcome may not
be observed by humans and hence can be optimized.

@@

\begin{figure}
\centering
\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/Mozilla477564.js}}
\caption{A cross-loop redundant bug in Mozilla }
\label{fig:Mozilla477564}
\end{figure}

@@

\paragraph{Coverage}
As shown in Table \ref{tab:root}, 
our taxonomy does cover all inefficient loops under study. 
Resultless loops are about as common as redundant loops
(24 vs. 21).
Intuitively, 0* loops, where no iteration produces any results,
are rare in mature software. In fact, no bugs in this
benchmark suite belong to this category.
All other root-cause sub-categories are well represented.

>>>>

\subsubsection{Resultless loops}
A resultless loop spends a lot of time in computation without
producing results useful after the loop.
It is further categorized based on which loop iterations are
(not) producing useful results.

@@

\begin{figure}
\centering
\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true,numbers=left}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/Mozilla347306.c}}
\vspace{-0.05in}
\caption{A resultless 0*1? bug in Mozilla}
\label{fig:Mozilla347306}
\end{figure}

{\textbf{0*1?}}:
This type of loops only produce results in the last iteration, if any. 
They are often related to search: check a sequence of elements one
by one until the right one is found.
Clearly, whether these loops are efficient or not depends on the
workload. When indeed inefficient, they are often
fixed by data-structure changes.
An example is shown in Figure \ref{fig:Mozilla347306}.
Large JavaScript files often fill the \texttt{script} list with 
tens of thousands of nodes and cause poor performance.
The patch simply replaced the \texttt{script} list with a hash table.
% This improve the performance by 52.60{\bf X}.

@@

\begin{figure}
\centering
\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/GCC46401.c}}
\vspace{-0.05in}
\caption{A resultless [0$|$1]* bug in GCC }
\label{fig:GCC46401}
\end{figure}

{\textbf{[0$|$1]*}}:
Each iteration in this type of loops may or may not produce results.
For some workload, the majority of iterations do not produce
results and cause performance problems perceived by users.
These problems are often solved
by adding a simple check to conditionally skip the expensive loop.
Figure~\ref{fig:GCC46401} shows such an example.
%how users perceive this problem
%explain how every iteration may or may not produce any results
%explain the fix
Users complained that compilation became extremely slow when
the \texttt{-Wsequence-point} checking is enabled.
The slowness was caused
by the \texttt{for} loop in the figure. As the algorithm
behind this loop has quadratic complexity in the number of operands in an expression, 
programs with long expressions suffer severe slow-downs.
After further diagnosis, developers observed that this loop rarely
had any side-effects, as the \texttt{if} condition was rarely satisfied.
At the end, the patch greatly improved performance through a
simple checking that allows the loop to be
skipped for most expressions.

@@

{\textbf{1*}}:
Loops in this category always generate results in almost all iterations. 
They are inefficient because their results are useless due to
some high-level semantic reasons.
Understanding and fixing this type of inefficiency problems often require
deep understanding of the program and are difficult to automate.
For example, several Mozilla performance problems are caused by 
loops that contain intensive GUI operations whose graphical outcome may not
be observed by humans and hence can be optimized.

@@

\begin{figure}
\centering
\lstset{basicstyle=\ttfamily\fontsize{8}{8}\selectfont,
     morekeywords={+},keepspaces=true}
  \mbox{\lstinputlisting[mathescape,boxpos=t]{figures/Mozilla477564.js}}
\vspace{-0.05in}
\caption{A cross-loop redundant bug in Mozilla }
\label{fig:Mozilla477564}
\end{figure}

@@

\paragraph{Coverage}
As shown in Table \ref{tab:root}, 
our taxonomy does cover all inefficient loops under study. 
Resultless loops are about as common as redundant loops
(24 vs. 21).
Not surprisingly, 0* loops
are rare in mature software. In fact, no bugs in this
benchmark suite belong to this category.
All other root-cause sub-categories are well represented.

### 2015-11-20 section/2_workless.tex
finish a pass of the whole paper

\Tool first identifies side-effect instructions that write to
variables defined outside the loop. 
\Tool analyzes all functions called
by a loop directly or indirectly --- a function $F$ that updates variables
defined outside $F$ makes the corresponding call statement in $F$'s
caller a side-effect instruction.
We consider all calls to library functions or through function pointers 
as having side effects, 
unless the library functions are specially marked by us in a white list.

@@

For a 0*1? loop, since it only generates results in the last iteration, we 
only need to know the total number of loop iterations 
to figure out the 
\textit{resultless rate} of the loop. The implementation is straightforward
--- we initialize a local counter to be 0 in the pre-header of the loop; we 
increase the counter by 1 in the loop header to count the number of 
iterations; we dump that counter to log when the loop exits.

>>>>

\Tool considers two types of instructions as side-effect instructions:
(1) write to heap or global variables;
(2) write to a stack variable, defined outside the loop, and the update value
may be used after the loop. The second condition is checked through liveness
analysis.
\Tool also analyzes all functions called
by a loop directly or indirectly --- a function $F$ that updates variables
defined outside $F$ makes the corresponding call statement in $F$'s
caller a side-effect instruction.
We consider all calls to library functions or through function pointers 
as having side effects, 
unless the library functions are specially marked by us in a white list.
This analysis is complete, but not sound. We may identify side-effect
instructions that are actually not.

@@

For a 0*1? loop, since it only generates results in the last iteration, we 
only need to know the total number of loop iterations 
to figure out the loop
\textit{resultful rate}, which we define as the ratio between
the number of iterations with
side effects and the total number of iterations. 
The implementation is straightforward
--- we initialize a local counter to be 0 in the pre-header of the loop; we 
increase the counter by 1 in the loop header to count the number of 
iterations; we dump that counter to log when the loop exits.

### 2015-11-20 section/3_redundancy.tex
finish a pass of the whole paper

Specifically, \Tool first identifies side-effect instructions in $C$.
%, as 
%discussed in Section \ref{sec:s_workless}. 
It then conducts 
backward static slicing 
from these instructions, considering both control and data
dependency. For every memory read instruction $r$ 
(\texttt{read(v)}) that static slicing encounters, \Tool checks whether $r$
satisfies either one of the following two conditions. 
If $r$ does, it is marked as
an \textit{input} instruction; otherwise, \Tool continues growing the
slice beyond $r$.
The two conditions are:
(1) the value of $v$ is defined outside $C$;
(2) $v$ is a heap or global variable.
The rationale for the first condition is that slicing outside $C$ is
unnecessary for redundancy judgement among instances of $C$.
The rationale for the second condition is that tracking data-dependency
through heap or global variables is complicated in multi-threaded C/C++
programs.

@@

\paragraph{Example}
Figure \ref{fig:Apache34464} shows a simplified code-snippet from Apache,
containing a cross-loop redundancy bug. 
The \texttt{for} loop starting on line 7 
searches a string \texttt{source} for a target sub-string 
\texttt{target}. Since the outer loop, which starts on line 2, appends one 
character to \texttt{source} in every iteration (line 30), 
every execution of the \texttt{for} loop is working on a similar 
\texttt{source} string from its previous execution, with a lot of redundancy. 
%applied to many \texttt{source} strings slightly different from each other
%with a lot of
%redundant computation.
%We use it to demonstrate how we identify \texttt{input}s below.

@@

\paragraph{Cross-iteration redundancy analysis}
Our high-level sampling strategy is straightforward:
randomly decide at the
beginning of every iteration whether to track the values returned by
source instructions in this iteration.

@@

First, we identify and avoid monitoring memory reads whose return values can be 
statically proved to
not change throughout one loop instance
or multiple loop instances. 
Since we implement \Tool at LLVM byte-code level, the majority of this
analysis is already done by LLVM, which lifts loop-invariant memory
accesses out of loops. 
For example, the read of \texttt{max} on line 12 of 
Figure \ref{fig:Apache34464} is an \textit{input} instruction of the 
\texttt{while} loop on the same line. LLVM identifies it as a loop invariant
and lifts the read of \texttt{max}
out of the loop. Consequently, \Tool only records
its value once during each loop, not each iteration.
However, LLVM does not apply its loop-invariant analysis to heap and global
variables, so \Tool conducts some extra analysis --- it identifies those
memory reads \texttt{read(v)}, where the address of \texttt{v} 
is a loop invariant and no instructions 
inside the loop can possibly write to that address. 
For example, \Tool can identify that the values of
\texttt{aNode->localName} and \texttt{aNode->namespaceURI} do not change
throughout one loop instance
in Figure~\ref{fig:Mozilla477564}. This analysis is sound but not complete.
\Tool may miss some loop invariants.

Second, we leverage the scalar evolution analysis in LLVM to remove the 
monitoring to some loop-induction related variables.
The scalar evolution analysis can tell which variables are
loop-induction variables (e.g., \texttt{i} on line 10
of Figure~\ref{fig:Apache34464}) and what are their strides. 
We then use this information in two ways. 
One is for cross-iteration redundancy checking.
If a loop iteration's \texttt{input} set contains a loop-induction
variable, we know different iterations' inputs would be different and 
hence conclude that there is no cross-iteration redundancy without any
run-time analysis. 
The other is for cross-loop redundancy checking.
When the address of a memory read is a loop induction variable, such as
\texttt{source[i]} on line 10 of Figure~\ref{fig:Apache34464},
we only record the address range at run time (i.e., the start and the
end), instead of every value returned by the memory read.
Of course, this optimization could lead to false positives, reporting
non-redundant loop computation as redundant.
This may happen when the corresponding
memory locations are modified across different loop instances.
In that case, although different loop instances work on variables read from
similar memory locations, the values they read are different.

>>>>

Specifically, \Tool first identifies side-effect instructions in $C$,
as 
discussed in Section \ref{sec:s_workless}. 
It then conducts 
backward static slicing 
from these instructions, considering both control and data
dependency. For every memory read instruction $r$ 
(\texttt{read(v)}) that static slicing encounters, \Tool checks whether $r$
satisfies either one of the following two conditions. 
If $r$ does, it is marked as
an \textit{input} instruction; otherwise, \Tool continues growing the
slice beyond $r$.
The two conditions are:
(1) the value of $v$ is defined outside $C$;
(2) $v$ is a heap or global variable.
The rationale for the first condition is that slicing outside $C$ is
unnecessary for redundancy judgement among instances of $C$.
The rationale for the second condition is that tracking data-dependency
through heap or global variables is complicated in multi-threaded C/C++
programs.

@@

\paragraph{Example}
Figure \ref{fig:Apache34464} shows a simplified code-snippet from Apache,
containing a cross-loop redundancy bug. 
The \texttt{for} loop starting on line 7 
searches a string \texttt{source} for a target sub-string 
\texttt{target}. Since the outer loop, which starts on line 2, appends one 
character to \texttt{source} in every iteration (line 30), 
every execution of the \texttt{for} loop is working on a similar 
\texttt{source} string from its previous execution, with a lot of redundancy.

@@

\paragraph{Cross-iteration redundancy analysis}
Our high-level sampling strategy is straightforward:
randomly decide at the
beginning of every iteration whether to track the values returned by
input instructions in this iteration.

@@

First, we avoid tracking multiple read instructions that we can statically 
prove to return the same value.
Since we implement \Tool in LLVM, we leverage the SSA construction and the
\textit{mem2reg} pass in LLVM to avoid unnecessary tracking of
stack variables.
For example, the read of \texttt{max} on line 12 of 
Figure \ref{fig:Apache34464} is an \textit{input} instruction of the 
\texttt{while} loop on the same line. LLVM identifies it as a loop invariant
and lifts the read of \texttt{max}
out of the loop. Consequently, \Tool only records
its value once during each loop, instead of each iteration.
LLVM is conservative in lifting heap/global variables out of
a loop, for fear of changing the location of potential exceptions, etc.
\Tool conducts a best-effort
loop invariant analysis for heap/global variables.
For example, \Tool identifies that the values of
\texttt{aNode->localName} and \texttt{aNode->namespaceURI} do not change
throughout one loop instance
in Figure~\ref{fig:Mozilla477564}, and hence only traces them once
outside the loop. This analysis is sound but not complete.
\Tool may miss some loop invariants. For in-house debugging, since we know
the problem-triggering inputs do not lead to exceptions, this optimization
is safe. For production-run usage, users can turn off this heap/global
variable related optimization.

Second, we leverage the scalar evolution analysis in LLVM to remove the 
monitoring to some loop-induction related variables.
The scalar evolution analysis can tell which variables are
loop-induction variables (e.g., \texttt{i} on line 10
of Figure~\ref{fig:Apache34464}) and what are their strides. 
We then use this information in two ways. 
One is for cross-iteration redundancy checking.
If a loop iteration's input set contains a loop-induction
variable, we know different iterations' inputs would be different and 
hence conclude that there is no cross-iteration redundancy without any
run-time analysis. 
The other is for cross-loop redundancy checking.
When the address of a memory read is a loop induction variable, such as
\texttt{source[i]} on line 10 of Figure~\ref{fig:Apache34464},
we only record the address range at run time (i.e., the start and the
end), instead of every value returned by the memory read.
This optimization could lead to false positives ---
different loop instances may work on variables read from
similar memory locations but with different values.
We did not encounter such false positives in our experiments.
